---
title: "Oxygen Analysis & Figures - Biogeochem. 2020"
author: "Emily Lacroix"
date: "8/28/2020"
output: 
  github_document:
    toc: true
  
---

```{r knitr setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 300)
```

## Set-Up

### Load libraries
```{r Libraries, results='hide', message=FALSE, warning=FALSE}

library(MASS)
library(MuMIn)
library(car)
library(readxl)
library(modelr)
library(tidyverse)

conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")

```

### Designate filepaths
```{r Input files}

data_file <- "~/Desktop/TempeCell_MasterSpreadsheet.xlsx"

```


###  Figure labels

```{r Field Key, warning=FALSE}
#Field key
fields <- 
  tribble(
    ~crop_type, ~fields,
    "Native Grassland", c("7-1", "8-6", "3-6"),
    "Riparian Grassland", c("RG-1", "RG-2", "RG-3"),
    "Wheat/Cover", c("1-6", "3-8", "7-3"),
    "Wheat/Fallow", c("1-7", "4-9", "8-3"),
  ) %>% 
  unnest()
```

```{r Moisture Labels}
moisture_label <- c("low" = "< 80 % WFPS", "high" = "> 80 % WFPS")
```


## Import & Clean Data

### Import  data
Filter out any Fe cores as they were not utilized in our incubations
```{r Gravimetric Data}
grav_data <- 
  read_xlsx(
    data_file, 
    sheet = "GravimetricData",
    na = "NA"
  ) %>%   
  filter(!str_detect(sample, "Fe"))
```


```{r DO data}
raw_do <- 
  read_xlsx(
    data_file, 
    sheet = "DissolvedOxygen",
    na = "NA"
  )
```


```{r Gas filled Bags}
gas_filled <- 
  read_xlsx(data_file, sheet = "GasFilledBags", na = "NA")
```

### Clean-up data

This code:

* sets all negative DO values to zero
* sets all DO values greater than 283 to 283
* omits any data points with a dilution factor greater than 1
* removes all data points where the bags filled with nitrogen gas
* classifies core moistures into two classes: above and below 80% WFPS
* calculates the oxygen consumption in each core
* classifies our data into different disturbance classes

```{r Clean Up DO}

do <- 
  raw_do %>% 
  mutate_at(vars(adj_do), ~ if_else(. < 0, 0, .)) %>% 
  mutate_at(vars(adj_do), ~ if_else(. > 283, 283, .)) %>% 
  mutate_at(vars(incubated_days, pressure), as.integer) %>% 
  left_join(fields, by = c("field" = "fields")) %>% 
  filter(df <= 1) %>% 
  anti_join(
    gas_filled, 
    by = 
      c(
        "field", 
        "incubated_days" = "day", 
        "pressure", 
        "moisture", 
        "disturbed"
      )
  ) %>% 
  left_join(
    grav_data %>% select(-sample), 
    by = 
      c("field", "incubated_days" = "day", "disturbed", "moisture")
  ) %>% 
  mutate(
    AMC = case_when(
      wfps < 0.80 ~ "low",
      wfps > 0.80 ~ "high")
  ) %>% 
  mutate_at(
    vars(AMC),
    ~ factor(., levels = c("low", "high"))
  ) %>% 
  mutate(
    total_do = adj_do * flush_vol / 1000,
    theoretical_max = 283 * flush_vol / 1000,
    umol_consumed = theoretical_max - total_do
  ) %>% 
  mutate(
    tillage = 
      case_when(
        crop_type %in% c("Wheat/Fallow", "Wheat/Cover") ~ "Tilled Cropland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == FALSE ~ "Undisturbed Grassland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == TRUE ~ "Sieved & Re-packed Grassland",
      )
  ) %>% 
  drop_na(pressure)

```

## Figures & Tables

### Table 3: Oxygen Consumed 
```{r Table 3 Mean Oxygen consumption}

do %>% 
  drop_na(umol_consumed) %>% 
  group_by(pressure, tillage) %>% 
  summarise(
    mean_umol_consumed = mean(umol_consumed, na.rm = TRUE),
    se_umol_consumed = sd(umol_consumed, na.rm = TRUE)/sqrt(n())
  ) %>% 
  arrange(tillage, pressure) %>%
  mutate_at(
    vars(mean_umol_consumed, se_umol_consumed), 
    ~ round(., digits = 3)
  ) %>% 
  knitr::kable()
  
```

### Table 3: Proportion of Oxygen Consumed
```{r Table 3 Proportion Oxygen Consumed}

do %>% 
  drop_na(umol_consumed) %>% 
  pivot_wider(
    id_cols = 
      c(
        "field", 
        "incubated_days", 
        "crop_type", 
        "wfps", 
        "AMC", 
        "tillage",
        "disturbed"
      ),
    names_from = "pressure",
    values_from = "umol_consumed"
  ) %>% 
  rename_at(
    vars(`100`,`300`,`500`),
    ~ paste0("umol_consumed_",.)
  ) %>% 
  mutate(
    total_umol_consumed = 
      umol_consumed_100 + umol_consumed_300 + umol_consumed_500,
    prop_consumed_100 = umol_consumed_100 / total_umol_consumed,
    prop_consumed_300 = umol_consumed_300 / total_umol_consumed,  
    prop_consumed_500 = umol_consumed_500 / total_umol_consumed,
    
  ) %>% 
  drop_na(total_umol_consumed) %>% 
  group_by(tillage) %>% 
  summarize(
    mean_prop_100 = mean(prop_consumed_100, na.rm = TRUE),
    se_prop_100 = sd(prop_consumed_100, na.rm = TRUE) / sqrt(n()),
    mean_prop_300 = mean(prop_consumed_300, na.rm = TRUE),
    se_prop_300 = sd(prop_consumed_300, na.rm = TRUE) / sqrt(n()),
    mean_prop_500 = mean(prop_consumed_500, na.rm = TRUE),
    se_prop_500 = sd(prop_consumed_500, na.rm = TRUE) / sqrt(n())
  ) %>% 
  pivot_longer(
    cols = c(
      mean_prop_100, 
      mean_prop_300, 
      mean_prop_500, 
      se_prop_100, 
      se_prop_300,
      se_prop_500
    )
  ) %>% 
  rename(prop_consumed = value) %>% 
  mutate(
    pressure = str_extract(name, "\\d{3}"),
    value_type = str_extract(name, "\\w{2,4}(?=_)")
  ) %>% 
  pivot_wider(
    id_cols = c("tillage", "pressure"),
    names_from = c("value_type"),
    values_from = "prop_consumed"
  ) %>% 
  mutate_at(vars(mean, se), round, digits = 4) %>% 
  knitr::kable()

```

### Figure S4: Boxplot of DO by disturbance and moisture

#### Calculate mean DO by pore, separated by moisture class

```{r DO means by moisture and pore size}

average_do_by_pore_moisture <- 
  do %>% 
  mutate(
    tillage = 
      case_when(
        crop_type %in% c("Wheat/Fallow", "Wheat/Cover") ~ "Tilled Cropland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == FALSE ~ "Undisturbed Grassland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == TRUE ~ "Sieved & Re-packed Grassland",
      )
  ) %>% 
  mutate_at(
    vars(tillage),
    ~ factor(
      ., 
      levels = 
        c(
          "Undisturbed Grassland", 
          "Tilled Cropland", 
          "Sieved & Re-packed Grassland" 
        ),
      labels = 
        c(
          "Undisturbed \nGrassland", 
          "Tilled \nCropland",
          "Sieved \nGrassland"
        )
      )
  ) %>% 
  drop_na(pressure) %>% 
  mutate_at(
    vars(pressure),
    ~ factor(
      .,
      levels = c(100, 300, 500),
      ordered = TRUE
    )
  ) %>%
  group_by(pressure, AMC) %>% 
  summarise(
    mean_do = mean(adj_do, na.rm = TRUE),
    se_do = sd(adj_do, na.rm = TRUE)/sqrt(n())
  ) 


```

```{r Fig S4}
do %>% 
  mutate(
    tillage = 
      case_when(
        crop_type %in% c("Wheat/Fallow", "Wheat/Cover") ~ "Tilled Cropland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == FALSE ~ "Undisturbed Grassland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == TRUE ~ "Sieved & Re-packed Grassland",
      )
  ) %>% 
  mutate_at(
    vars(tillage),
    ~ factor(
      ., 
      levels = 
        c(
          "Undisturbed Grassland", 
          "Tilled Cropland", 
          "Sieved & Re-packed Grassland" 
        ),
      labels = 
        c(
          "Undisturbed Grassland", 
          "Tilled Cropland",
          "Sieved Grassland"
        )
      )
  ) %>% 
  drop_na(pressure) %>% 
  mutate_at(
    vars(pressure),
    ~ factor(
      .,
      levels = c(100, 300, 500), 
      ordered = TRUE 
    )
  ) %>% 
  ggplot(aes(x = pressure, y = adj_do)) + 
  geom_hline(aes(yintercept = 283), linetype = 2, color = "darkgrey") +
  geom_hline(aes(yintercept = 15), linetype = 2, color = "darkgrey") +
  geom_boxplot(
    aes(fill = tillage), 
    position = position_dodge(0.8),
    color = "black"
  ) +
  geom_pointrange(
    data = average_do_by_pore_moisture,
    aes(
      x = pressure,
      y = mean_do,
      ymin = mean_do - se_do,
      ymax = mean_do + se_do
    ),
    color = "darkred",
    alpha = 0.8
  ) +
  scale_fill_brewer(palette = "Greys") + 
  scale_x_discrete(
    labels = 
      c(
        expression(paste("> 3", mu, "m")),
        expression(paste("1 - 3", mu, "m")),
        expression(paste("0.6 - 1", mu, "m"))
      )
  ) +
  facet_grid(
    rows = vars(AMC),
    labeller = 
      labeller(
        AMC = moisture_label,
        pressure = label_parsed
      )
  ) +
  theme_bw() + 
  theme(
    legend.position = "none",
    axis.ticks.x = element_line(color = "grey92"),
    aspect.ratio = 0.5,
    panel.grid = element_blank()
  ) + 
  labs(
    fill = "Degree of \nDisturbance",
    y = expression(paste("Dissolved Oxygen (",mu,"M)")),
    x = "Effective Pore Diameter"
  ) 

```


### Figure 3: Boxplot of DO by disturbance, averaged across moisture classes

#### Calculate the means for each pore size

```{r DO means by pore size}
average_do_by_pore <- 
  do %>% 
  mutate(
    tillage = 
      case_when(
        crop_type %in% c("Wheat/Fallow", "Wheat/Cover") ~ "Tilled Cropland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == FALSE ~ "Undisturbed Grassland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == TRUE ~ "Sieved & Re-packed Grassland",
      )
  ) %>% 
  mutate_at(
    vars(tillage),
    ~ factor(
      ., 
      levels = 
        c(
          "Undisturbed Grassland", 
          "Tilled Cropland", 
          "Sieved & Re-packed Grassland" 
        ),
      labels = 
        c(
          "Undisturbed \nGrassland", 
          "Tilled \nCropland",
          "Sieved \nGrassland"
        )
      )
  ) %>% 
  drop_na(pressure) %>% 
  mutate_at(
    vars(pressure),
    ~ factor(
      .,
      levels = c(100, 300, 500),
      ordered = TRUE
    )
  ) %>%
  group_by(pressure) %>% 
  summarise(
    mean_do = mean(adj_do, na.rm = TRUE),
    se_do = sd(adj_do, na.rm = TRUE)/sqrt(n())
  ) 

```

```{r}
average_do_by_pore %>% knitr::kable()
```


__Plot__
```{r Fig 3}
do %>% 
  mutate(
    tillage = 
      case_when(
        crop_type %in% c("Wheat/Fallow", "Wheat/Cover") ~ "Tilled Cropland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == FALSE ~ "Undisturbed Grassland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == TRUE ~ "Sieved & Re-packed Grassland",
      )
  ) %>% 
  mutate_at(
    vars(tillage),
    ~ factor(
      ., 
      levels = 
        c(
          "Undisturbed Grassland", 
          "Tilled Cropland", 
          "Sieved & Re-packed Grassland" 
        ),
      labels = 
        c(
          "Undisturbed Grassland", 
          "Tilled Cropland",
          "Disturbed Grassland"
        )
      )
  ) %>% 
  drop_na(pressure) %>% 
  mutate_at(
    vars(pressure),
    ~ factor(
      .,
      levels = c(100, 300, 500), 
      ordered = TRUE 
    )
  ) %>% 
  ggplot() + 
  geom_hline(aes(yintercept = 283), linetype = 2, color = "black") +
  geom_hline(aes(yintercept = 15), linetype = 2, color = "black") +
  geom_boxplot(
    aes(x = pressure, y = adj_do, fill = tillage), 
    position = position_dodge(0.8),
  ) + 
  geom_pointrange(
    data = average_do_by_pore,
    aes(
      x = pressure,
      y = mean_do,
      ymin = mean_do - se_do,
      ymax = mean_do + se_do
    ),
    color = "darkred",
    alpha = 0.8
  ) +
  scale_fill_brewer(palette = "Greys") + 
  scale_x_discrete(
    labels =
      c(
        expression(paste("> 3", mu, "m")),
        expression(paste("1 - 3", mu, "m")),
        expression(paste("0.6 - 1", mu, "m"))
      )
  ) +
  theme_bw() + 
  theme(
    rect = element_rect(fill = "transparent"),
    legend.position = "right",
    legend.text = element_text(margin = margin(b = 2, t = 2, unit = "pt")),
    text = element_text(family = "Arial", color = "black"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "transparent"), 
    plot.background = element_rect(fill = "transparent", color = NA)
  ) + 
  labs(
    fill = NULL,
    y = expression(paste("Dissolved Oxygen (",mu,"M)")),
    x = "Effective Pore Diameter"
  ) 



```

### Figure 4: DO concentration over time
```{r Figure 4}

do %>% 
  filter(disturbed == FALSE) %>% 
  drop_na(pressure) %>% 
  mutate_at(
    vars(pressure),
    ~ factor(
      .,
      levels = c(100, 300, 500), 
      ordered = TRUE, 
      labels = 
        c(
          expression(paste("> 3", mu, "m")),
          expression(paste("1 - 3", mu, "m")),
          expression(paste("0.6 - 1", mu, "m"))
        )
    )
  ) %>% 
  group_by(incubated_days, pressure, AMC) %>% 
  summarise(
    adj_do_mean = mean(adj_do, na.rm = TRUE),
    se_adj_do = sd(adj_do, na.rm = TRUE) / sqrt(n())
  ) %>% 
  ggplot(
    aes(
      x = incubated_days, 
      color = AMC, 
      fill = AMC, 
      y = adj_do_mean
    )
  ) + 
  geom_point(size = 2, shape = 21) +
  geom_errorbar(
    aes(
      ymin = adj_do_mean - se_adj_do, 
      ymax = adj_do_mean + se_adj_do
    ),
    width = 1
  ) +
  geom_line(aes(linetype = AMC)) + 
  geom_hline(yintercept = 283, linetype = 2) + 
  scale_x_continuous(breaks = c(7, 14, 28)) + 
  scale_color_manual(values = c("gray60", "grey12")) + 
  scale_fill_manual(
    values = c("gray60", "grey12"), 
    labels = c("< 80", "> 80")
  ) + 
  scale_linetype_manual(
    values = c(2,3)
  ) + 
  facet_grid(
    cols = vars(pressure),
    labeller = label_parsed
  ) + 
  labs(
    y = expression(paste("Dissolved Oxygen (",mu,"M)")),
    x = "Days Incubated",
    fill = " Moisture \n(% WFPS)"
  ) + 
  guides(
    color = "none",
    linetype = "none"
  ) +
  theme_bw() + 
  theme(
    legend.title = element_text(size = 8, face = "bold", color = "black"),
  ) 

```


## Statistics 

### Sample Sizes (N)
```{r Sample Size Table}
do %>%
  drop_na(umol_consumed) %>%
  group_by(pressure, tillage, incubated_days) %>% 
  summarise(
    n = n()
  ) %>% 
  knitr::kable()
```

### Checking Normality for adj_do values
```{r Shapiro Test for Normality - no transformation}
do %>% 
  drop_na(adj_do) %>% 
  group_by(tillage, pressure) %>% 
  summarise(
    normality_stat = 
      if_else(
        sd(adj_do, na.rm = TRUE) != 0, 
        shapiro.test(adj_do)$statistic, 
        NA_real_
      ),
    normality_p_value =  
      if_else(
        sd(adj_do, na.rm = TRUE) != 0, 
        shapiro.test(adj_do)$p.value, 
        NA_real_
      )
  ) %>% 
  knitr::kable()
```

The data are not normal.

### Check for equal variances
```{r}

leveneTest(
  adj_do ~ tillage*pressure, 
  data = do %>% 
  mutate_at(vars(pressure), as.factor)
)
```
The variances amongst the tillage x pressure groups are unequal. 

### Pairwise Wilcoxon Rank Sum Test

#### Table S5: DO Across Pore Sizes 

```{r Table S5}

pairwise.wilcox.test(
  x = do$adj_do,
  g = do$pressure,
  p.adjust.method = "BH",
  paired = FALSE
)

```

NOTE - result is the same when analysis is performed when re-doing the analysis on only a subset of the samples that had paired observations across pore sizes (paired = TRUE). See below:

```{r}

paired_do <-
  do %>% 
  pivot_wider(
    id_cols = c(field, incubated_days, moisture, disturbed),
    names_from = c(pressure),
    values_from = c(adj_do)
  ) %>% 
  filter_at(vars(`100`, `300`, `500`), all_vars(!is.na(.))) %>% 
  pivot_longer(
    cols = c(`100`, `300`, `500`), 
    names_to = "pressure", 
    values_to = "adj_do"
  )

pairwise.wilcox.test(
  x = paired_do$adj_do,
  g = paired_do$pressure,
  p.adjust.method = "BH",
  paired = TRUE
)
 
```


#### Table S5: DO Across Disturbance Treatments by Pore Size

__Largest pores__
```{r Largest pores}
largest_pore_do <- 
  do %>% 
  filter(pressure == 100)

pairwise.wilcox.test(
  x = largest_pore_do$adj_do,
  g = largest_pore_do$tillage,
  p.adjust.method = "BH",
  paired = FALSE
)

```


__Medial pores__

```{r Medium pores}
medium_pore_do <- 
  do %>% 
  filter(pressure == 300)

pairwise.wilcox.test(
  x = medium_pore_do$adj_do,
  g = medium_pore_do$tillage,
  p.adjust.method = "BH",
  paired = FALSE
)

```


__Smallest pores__
```{r Smallest pores}
smallest_pore_do <- 
  do %>% 
  filter(pressure == 500)

pairwise.wilcox.test(
  x = smallest_pore_do$adj_do,
  g = smallest_pore_do$tillage,
  p.adjust.method = "BH",
  paired = FALSE
)

```

#### Table S7: Across Pore Sizes, Separated by Moisture Class

__Table S6: <80% WFPS__

```{r Table S7, Drier Soils}

drier_pore_do <- 
  do %>% 
  filter(AMC == "low")

pairwise.wilcox.test(
  x = drier_pore_do$adj_do,
  g = drier_pore_do$pressure,
  p.adjust.method = "BH",
  paired = FALSE
)

```

__Table S6: >80% WFPS__

```{r Table S7, Wetter Soils}
wetter_pore_do <- 
  do %>% 
  filter(AMC == "high")

pairwise.wilcox.test(
  x = wetter_pore_do$adj_do,
  g = wetter_pore_do$pressure,
  p.adjust.method = "BH",
  paired = FALSE
)

```


### GLMM fits 

Create a dataframe with unique observation identifiers to use in GLMMs
```{r make do_core_id}

do_core_id <-
  do %>% 
  group_by(field, incubated_days, disturbed, moisture) %>% 
  pivot_wider(
    .,
    id_cols = c(
      field, 
      incubated_days, 
      disturbed, 
      moisture, 
      AMC,
      crop_type,
      bulk_dens,
      wfps,
      dry_mass,
      tillage
    ),
    names_from = pressure, 
    values_from = adj_do
  ) %>% 
  rowid_to_column() %>% 
  pivot_longer(
    cols = c(`100`, `300`, `500`), 
    names_to = "pressure", 
    values_to = "adj_do"
  ) %>% 
  rename(
    core_id = rowid
  ) %>% 
  rowid_to_column(var = "obsv_id")

```

#### GLMM for overall
```{r Table 2: glmm overall, results="hide", message=FALSE}

glmmPQL_overall <-
  glmmPQL(
    adj_do ~ tillage + bulk_dens * wfps + incubated_days + pressure,
    random = list(field = ~1, core_id = ~1),
    family = quasipoisson(link = "log"),
    data = 
      do_core_id %>% 
      ungroup() %>% 
      mutate_at(vars(adj_do), as.integer) %>% 
      mutate_at(vars(pressure, field), as.factor)
  )


```

```{r}
glmmPQL_overall %>% 
  summary()

```

#### GLMM for > 3.0 um
```{r Table 2: > 3 um, results="hide", message=FALSE}

glmmPQL_do100 <-
  glmmPQL(
    adj_do ~ tillage + bulk_dens * wfps + incubated_days,
    random = list(field = ~1),
    family = quasipoisson(link = "log"),
    data = 
      do_core_id %>% 
      ungroup() %>% 
      filter(pressure == 100) %>% 
      mutate_at(vars(adj_do), as.integer) %>% 
      mutate_at(vars(pressure, field), as.factor)
  )

```


```{r}
glmmPQL_do100 %>% summary() 
```

#### GLMM for 1.0 - 3.0 um
```{r glmmPQL 1-3 um, results="hide", message=FALSE}

glmmPQL_do300 <-
  glmmPQL(
    adj_do ~ tillage + bulk_dens * wfps + incubated_days,
    random = list(field = ~1),
    family = quasipoisson(link = "log"),
    data = 
      do_core_id %>% 
      ungroup() %>% 
      filter(pressure == 300) %>% 
      mutate_at(vars(adj_do), as.integer) %>% 
      mutate_at(vars(pressure, field), as.factor)
  )

```

```{r}
glmmPQL_do300 %>% summary()
```

#### GLMM for 0.6 - 1.0 um

```{r glmmPQL 0.6 to 1 um, results="hide", message=FALSE}

glmmPQL_do500 <-
  glmmPQL(
    adj_do ~ tillage + bulk_dens * wfps + incubated_days,
    random = list(field = ~1),
    family = quasipoisson(link = "log"),
    data = 
      do_core_id %>% 
      ungroup() %>% 
      filter(pressure == 500) %>% 
      mutate_at(vars(adj_do), as.integer) %>% 
      mutate_at(vars(pressure, field), as.factor)
  )

```

```{r}
glmmPQL_do500 %>% summary()
```

#### ESM GLMM - Interactions
```{r }
glmmPQL_esm <-
  glmmPQL(
    adj_do ~ tillage * pressure + bulk_dens * wfps + incubated_days,
    random = list(field = ~1, core_id = ~1),
    family = quasipoisson(link = "log"),
    data = 
      do_core_id %>% 
      ungroup() %>% 
      mutate_at(vars(adj_do), as.integer) %>% 
      mutate_at(vars(pressure, field), as.factor)
  )

```

```{r}
glmmPQL_esm %>% summary()

```

