---
title: "GHG Analysis & Figures"
author: "Emily Lacroix"
date: "4/16/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 300)
```


## Set-Up
### Load libraries
```{r,  results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)

conflicted::conflict_prefer("recode", "dplyr")
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
```

### Designate file paths

```{r}

data_file <- "~/Desktop/TempeCell_MasterSpreadsheet.xlsx"

```

### Figure labels

```{r}
moisture_label <- c("low" = "< 80 % WFPS", "high" = "> 80 % WFPS")
```

```{r}

field_labels <- 
  c(
    "RG-3" = "Riparian Grassland",
    "RG-2" = "Riparian Grassland",
    "RG-1" = "Riparian Grassland",
    "RG" = "Riparian Grassland",
    "3-8" = "Wheat/Cover",
    "1-6" = "Wheat/Cover",
    "7-3" = "Wheat/Cover",
    "4-9" = "Wheat/Fallow",
    "1-7" = "Wheat/Fallow",
    "8-3" = "Wheat/Fallow",
    "8-6" = "Native Grassland",
    "3-6" = "Native Grassland",
    "7-1" = "Native Grassland"
  )

```

### Set constants

```{r}
jar_vol_l <- 0.473

volume_one_mol <- 1 * 8.314 * 298.15 / 101.325   #Ideal gas law: V = nRT/P

mols_per_l <- 1 / volume_one_mol

molar_mass_c <- 12.01

molar_mass_co2 <- 12.01 + (2 * 16) 

core_surface_area_m2 <- 0.0001 * pi * (5.40 / 2)^2 

```

## Import & Clean Data

### Import GHG Data

```{r}
ghg <- 
  data_file %>% 
  read_excel(sheet = "CO2", na = "NA")

```


### Import gravimetric data 
Filter out any Fe cores as they were not utilized in our incubations
```{r}

grav_data <- 
  read_xlsx(
    data_file, 
    sheet = "GravimetricData",
    na = "NA"
  ) %>% 
  filter(!str_detect(sample, "Fe"))

```

### Import TC data
Summarize OC data to the field-averaged level
```{r}

tc <- 
  data_file %>% 
  read_excel(sheet = "TotalCarbon", na = "NA") %>% 
  group_by(field) %>% 
  summarise(
    mean_perc_c = mean(perc_c, na.rm = TRUE), 
    mean_perc_n = mean(perc_n, na.rm = TRUE)
  ) 

```

## Flux Calculations

### Calculate the normalized fluxes 

This code:

* joins the CO2, gravimetric, and TC data together
* classifies core moistures into two classes: above and below 80% WFPS
* calculates the amount of CO2 produced in each jar over 24 hours
* normalizes this amount to the g of soil C in that core
* converts the normalized amount to a flux rate 
* summarizes the flux data by moisture, disturbance, and incubation day

```{r}

ghg_means_normalized <- 
  ghg %>% 
  left_join(
    grav_data %>% select(-sample), 
    by = c("field", "incubation_day" = "day", "disturbed", "moisture")
  ) %>% 
  left_join(tc, by = "field") %>% 
  mutate(
    AMC = case_when(
      wfps < 0.80 ~ "low",
      wfps >= 0.80 ~ "high")
  ) %>% 
  mutate_at(
    vars(AMC),
    ~ factor(., levels = c("low", "high"))
  ) %>% 
  mutate(
    g_soil_c = (mean_perc_c / 100) * dry_mass,
    headspace_vol_l = jar_vol_l - (ring_vol_cm3 / 1000),
    co2_accum_jar_umol = 
           10^6 * (co2 / 10^6) * headspace_vol_l * mols_per_l,
    co2_accum_jar_g = (co2_accum_jar_umol / 10^6) * molar_mass_co2,
    co2_accum_jar_umol_normalized = co2_accum_jar_umol / g_soil_c,
    co2_accum_g_norm = co2_accum_jar_umol_normalized / 10^6 * molar_mass_co2,
    co2_flux_g_m2_hr_norm = co2_accum_g_norm / (24 * core_surface_area_m2)
  ) %>% 
  mutate_at(
    vars(field),
    ~ recode(., !!!field_labels)
  ) %>% 
  mutate(
    tillage = 
      case_when(
        field %in% c("Wheat/Fallow", "Wheat/Cover") ~ "Tilled Cropland",
        field %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == FALSE ~ "Undisturbed Grassland",
        field %in% c("Native Grassland", "Riparian Grassland") & 
          disturbed == TRUE ~ "Disturbed Grassland",
      )
  ) %>% 
  mutate_at(
    vars(tillage),
    ~ factor(
      .,
      levels =
        c("Undisturbed Grassland", "Tilled Cropland","Disturbed Grassland"),
      labels = 
        c("Undisturbed\nGrassland", "Tilled\nCropland","Disturbed\nGrassland")
      )
  ) %>% 
  mutate_at(vars(day_sampled), as.integer) %>% 
  group_by(day_sampled, tillage, AMC) %>% 
  summarise( 
    mean_g_co2_rate = mean(co2_flux_g_m2_hr_norm, na.rm = TRUE),
    se_g_co2_rate = 
      sd(co2_flux_g_m2_hr_norm, na.rm = TRUE) / sqrt(n())
  ) 

ghg_means_normalized 

```

### Calculate cumulative CO2 production

Using the mean fluxes calculated above, this code:

* calculates the cumulative CO2 production by moisture and disturbance 
```{r}
cumulative_co2_produced <- 
  ghg_means_normalized %>% 
  pivot_wider(
    id_cols = c(tillage, AMC), 
    names_from = day_sampled, 
    values_from = c(mean_g_co2_rate, se_g_co2_rate)
  ) %>% 
  mutate(
    cumulative_g_co2_norm = 
      (mean_g_co2_rate_1 * 24) + 
      ((mean_g_co2_rate_1 + mean_g_co2_rate_3)/2 * 24 * 2) + 
      ((mean_g_co2_rate_3 + mean_g_co2_rate_7)/2 * 24 * 4) + 
      ((mean_g_co2_rate_7 + mean_g_co2_rate_14)/2 * 24 * 7) + 
      ((mean_g_co2_rate_14 + mean_g_co2_rate_28)/2 * 24 * 14),
    se_g_co2_norm = 
      (se_g_co2_rate_1 * 24) + 
      ((se_g_co2_rate_1 + se_g_co2_rate_3)/2 * 24 * 2) + 
      ((se_g_co2_rate_3 + se_g_co2_rate_7)/2 * 24 * 4) + 
      ((se_g_co2_rate_7 + se_g_co2_rate_14)/2 * 24 * 7) + 
      ((se_g_co2_rate_14 + se_g_co2_rate_28)/2 * 24 * 14) 
  ) %>% 
  select(tillage, AMC, cumulative_g_co2_norm, se_g_co2_norm)

cumulative_co2_produced
```

### Calculated averages reported in text

__Averaged by WFPS__
```{r}
cumulative_co2_produced %>% 
  group_by(AMC) %>% 
  summarise(
    avg_cumulative_g_co2_norm = mean(cumulative_g_co2_norm, na.rm = TRUE)
  ) %>% 
  knitr::kable()

```

__Averaged by disturbance__
```{r}
cumulative_co2_produced %>% 
  group_by(tillage) %>% 
  summarise(
    avg_cumulative_g_co2_norm = mean(cumulative_g_co2_norm, na.rm = TRUE)
  ) 
```

## Figures 

### Figure 2 - Cumulative CO2 by moisture & tillage
```{r}
cumulative_co2_produced %>% 
  ggplot(aes(x = tillage, y = cumulative_g_co2_norm)) + 
  geom_line(aes(group = AMC, linetype = AMC)) +
  geom_errorbar(
    aes(
      ymin = cumulative_g_co2_norm - se_g_co2_norm,
      ymax = cumulative_g_co2_norm + se_g_co2_norm,
      color = tillage
    ),
    width = 0.1
  ) + 
  geom_point(aes(fill = tillage), size = 5, shape = 21) + 
  scale_color_manual(values = c("gray60", "grey39", "grey12")) + 
  scale_fill_manual(values = c("gray60", "grey39", "grey12")) + 
  scale_y_continuous(limits = c(0, 160)) + 
  scale_linetype_manual(
    values = c(2,3),
    labels = moisture_label
  ) + 
  guides(
    fill = "none",
    color = "none"
  ) + 
  theme_bw() + 
  theme(
    panel.grid = element_line(colour = "grey94"),
    panel.grid.minor.x = element_blank(),
    legend.position = c(0.15, 0.85),
    axis.text.x = element_text(size = 16, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.title.y = element_text(size = 16), 
    legend.text = element_text(size = 14)
  ) + 
  labs(
    y = expression(Total~CO[2]~Flux~(g~CO[2]~m^-2~g~soil~C^-1)),
    x = NULL,
    color = NULL,
    linetype = NULL
  ) 

```

### Figure S4 - CO2 by moisture & disturbance over time
```{r}

ghg_means_normalized %>%   
  ggplot(
    aes(x = day_sampled, color = tillage, y = mean_g_co2_rate)
  ) +
  geom_line(linetype = 2) +
  geom_errorbar(
    aes(
      ymin = mean_g_co2_rate - se_g_co2_rate, 
      ymax = mean_g_co2_rate + se_g_co2_rate
    ),
    width = 0.8
  ) + 
  geom_point(aes(fill = tillage), size = 3, shape = 21, color = "black") +
  scale_x_continuous(breaks = c(1, 3, 7, 14, 28)) + 
  scale_color_manual(values = c("gray60", "grey39", "grey12")) + 
  scale_fill_manual(values = c("gray60", "grey39", "grey12")) + 
  guides(
    color = "none",
    fill = guide_legend(reverse = TRUE)
  ) + 
  facet_grid(
    cols = vars(AMC),
    labeller = labeller(AMC = moisture_label)
  ) + 
  theme_bw() + 
  theme(
    panel.grid = element_line(colour = "grey94"),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.title.y = element_text(size = 16), 
    axis.title.x = element_text(size = 16), 
    legend.text = element_text(size = 12),
    strip.text.x = element_text(size = 14, face = "bold"),
    legend.key.height =  unit(1.5, "cm")
  ) + 
  labs(
    y = expression(CO[2]~Flux~(g~CO[2]~m^-2~h^-1~g~soil~C^-1)),
    x = "Days Incubated",
    color = NULL,
    fill = NULL
  ) 

```



