---
title: "Ancillary Analyses"
author: "Emily Lacroix"
date: "8/21/2020"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 300)
```


## Set-Up
### Load libraries
```{r,  results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(car)

conflicted::conflict_prefer("recode", "dplyr")
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
```

### Designate file paths
```{r}

data_file <- "~/Desktop/TempeCell_MasterSpreadsheet.xlsx"

```

### Figure labels

```{r}

field_labels <- 
  c(
    "RG-3" = "Riparian Grassland",
    "RG-2" = "Riparian Grassland",
    "RG-1" = "Riparian Grassland",
    "RG" = "Riparian Grassland",
    "3-8" = "Wheat/Cover",
    "1-6" = "Wheat/Cover",
    "7-3" = "Wheat/Cover",
    "4-9" = "Wheat/Fallow",
    "1-7" = "Wheat/Fallow",
    "8-3" = "Wheat/Fallow",
    "8-6" = "Native Grassland",
    "3-6" = "Native Grassland",
    "7-1" = "Native Grassland"
  )

```


## Import Data

### TC and TN data
```{r}
tc <- 
  data_file %>% 
  read_excel(sheet = "TotalCarbon", na = "NA") 

```

### pH data
```{r}
ph <- 
  data_file %>% 
  read_excel(sheet = "pH")

```


### Gravimetric data
```{r}
grav_data <- 
  read_xlsx(
    data_file, 
    sheet = "GravimetricData",
    na = "NA"
  ) 

```

### Texture data
```{r}
texture <-
  data_file %>% 
  read_excel(sheet = "Texture")

```

### DOC data
```{r}
doc <-
  data_file %>% 
  read_excel(sheet = "DissolvedOrganicCarbon") %>% 
  select(-comments) %>% 
  filter(!str_detect(sample_name, "Blank")) %>% 
  mutate(
    disturbed = if_else(str_detect(sample_name, "-D"), TRUE, FALSE)
  )

```

## Figures & Analysis

### Table 1: % OC and TN 
```{r}
tc %>% 
  mutate_at(
    vars(field),
    ~ recode(., !!!field_labels)
  ) %>% 
  mutate(
    grass_crop = 
      if_else(
        field %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  ) %>% 
  group_by(grass_crop) %>% 
  summarise_at(
    vars(perc_c, perc_n), 
    list(
      mean = ~mean(., na.rm = TRUE), 
      se = ~sd(., na.rm = TRUE) / sqrt(n())
    )
  ) %>% 
  select(grass_crop, perc_c_mean, perc_c_se, perc_n_mean, perc_n_se) %>% 
  knitr::kable()

```

### OC & TN Stats
```{r}
tc_stats <- 
  tc %>% 
  mutate_at(
    vars(field),
    ~ recode(., !!!field_labels)
  ) %>% 
  mutate(
    grass_crop = 
      if_else(
        field %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  )
  
```

```{r}
tc_stats
```


```{r}
tc_stats %>% 
  group_by(grass_crop) %>% 
  ggplot(aes(x = perc_c, fill = grass_crop)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous() + 
  theme_bw()
```

```{r}
tc_stats %>% 
  group_by(grass_crop) %>% 
  ggplot(aes(x = perc_n, fill = grass_crop)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous() + 
  theme_bw()
```


```{r}

tc_stats %>% 
  group_by(grass_crop) %>% 
  summarise(
    normality_shapiro_p_oc = shapiro.test(perc_c)$p.value,
    normality_shapiro_p_tn = shapiro.test(perc_n)$p.value,
  )

```

```{r}
tc_stats %>% 
  group_by(grass_crop) %>% 
  summarise(
    normality_shapiro_p_oc = shapiro.test(sqrt(perc_c))$p.value,
    normality_shapiro_p_tn = shapiro.test(sqrt(perc_n))$p.value,
  )
```
```{r}
tc_stats %>% 
  group_by(grass_crop) %>% 
  summarise(
    normality_shapiro_p_oc = shapiro.test(perc_c^-3)$p.value,
    normality_shapiro_p_tn = shapiro.test(perc_n^-3)$p.value,
  )
```

```{r}
tc_stats %>% 
  group_by(grass_crop) %>% 
  summarise(
    normality_shapiro_p_oc = shapiro.test(log(perc_c))$p.value,
    normality_shapiro_p_tn = shapiro.test(log(perc_n))$p.value,
  )
```


The grassland data are not normally distributed. Log, sqrt, and cube root transofrmations still fail to meet the assumptions for normality. Checking for equal variances. 

```{r}
leveneTest(perc_c ~ grass_crop, data = tc_stats)
leveneTest(perc_n ~ grass_crop, data = tc_stats)
```
Variances are not equal. 

```{r}
wilcox.test(
  tc_stats$perc_c ~ tc_stats$grass_crop, 
  paired = FALSE
)
```

```{r}
wilcox.test(
  tc_stats$perc_n ~ tc_stats$grass_crop, 
  paired = FALSE
)
```


### Table 1: pH
```{r}

ph %>% 
  mutate(
    grass_crop = 
      if_else(
        crop %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  ) %>% 
  group_by(grass_crop) %>% 
  summarise_at(
    vars(pH), 
    list(
      mean = ~mean(., na.rm = TRUE), 
      se = ~sd(., na.rm = TRUE) / sqrt(n())
    )
  ) %>% 
  knitr::kable()
  
```

### pH Stats
```{r}
ph_stats <- 
  ph %>% 
  mutate(
    grass_crop = 
      if_else(
        crop %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  ) 
```


```{r}
ph_stats %>% 
  group_by(grass_crop) %>% 
  ggplot(aes(x = pH, fill = grass_crop)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous() + 
  theme_bw()

```

```{r}
ph_stats %>% 
  group_by(grass_crop) %>% 
  summarise(
    normality_p_val = shapiro.test(pH)$p.value
  )

```
The crop data are non-normal

```{r}
leveneTest(pH ~ grass_crop, data = ph_stats)
```
The variances are equal. Use Wilcoxon rank-sum (Mann Whitney U) test. 

```{r}
wilcox.test(
  ph_stats$pH ~ ph_stats$grass_crop, 
  paired = FALSE
)
```

### Table 1: Bulk Density
This focuses only on field cores, not disturbed cores as that bulk density is contrived. 
```{r}
grav_data %>% 
  mutate_at(
    vars(field),
    ~ recode(., !!!field_labels)
  ) %>% 
  filter(disturbed == FALSE) %>% 
  mutate(
    grass_crop = 
      if_else(
        field %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  ) %>% 
  group_by(grass_crop) %>% 
  summarise_at(
    vars(bulk_dens), 
    list(
      mean_bulk_dens = ~mean(., na.rm = TRUE), 
      se_bulk_dens = ~sd(., na.rm = TRUE) / sqrt(n())
    )
  ) %>% 
  knitr::kable()

```

### Bulk Density Stats
```{r}
bulk_dens_stats <- 
  grav_data %>% 
  mutate_at(
    vars(field),
    ~ recode(., !!!field_labels)
  ) %>% 
  filter(disturbed == FALSE) %>% 
  mutate(
    grass_crop = 
      if_else(
        field %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  ) 

```
```{r}
bulk_dens_stats %>% 
  group_by(grass_crop) %>% 
  ggplot(aes(x = bulk_dens, fill = grass_crop)) + 
  geom_histogram() +
  #geom_density(alpha = 0.5) + 
  scale_x_continuous() + 
  theme_bw()
```

Test for normality 
```{r}
bulk_dens_stats %>% 
  group_by(grass_crop) %>% 
  summarise(
    normality_p_val = shapiro.test(bulk_dens)$p.value
  )
```


Test for equal variances
```{r}
leveneTest(bulk_dens ~ grass_crop, data = bulk_dens_stats)
```

Variances are equal. 

Wilcoxon Rank Sum Test
```{r}
wilcox.test(
  bulk_dens_stats$bulk_dens ~ bulk_dens_stats$grass_crop, 
  paired = FALSE
)
```

### Table 1: Texture
```{r}
texture %>% 
  mutate(
    grass_crop = 
      case_when(
        str_detect(sample, "^(WF|WC|CW)") ~ "crop",
        str_detect(sample, "^(NG|RG)") ~ "grass"
      )
  ) %>% 
  group_by(grass_crop) %>% 
  summarise_at(
    vars(perc_sand, perc_silt, perc_clay), 
    list(
      mean = ~mean(., na.rm = TRUE),
      se = ~sd(., na.rm = TRUE) / sqrt(n())
    )
  ) %>% 
  select(
    grass_crop,
    perc_sand_mean,
    perc_sand_se,
    perc_silt_mean,
    perc_silt_se,
    perc_clay_mean,
    perc_clay_se
  ) %>% 
  knitr::kable()

```

### Texture Stats

```{r}
texture_stats <- 
  texture %>% 
    mutate(
    grass_crop = 
      case_when(
        str_detect(sample, "^(WF|WC|CW)") ~ "crop",
        str_detect(sample, "^(NG|RG)") ~ "grass"
      )
  )
```

```{r}
texture_stats %>%  
  group_by(grass_crop) %>% 
  ggplot(aes(x = perc_sand, fill = grass_crop)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous() + 
  theme_bw()

```

```{r}
texture %>% 
    mutate(
    grass_crop = 
      case_when(
        str_detect(sample, "^(WF|WC|CW)") ~ "crop",
        str_detect(sample, "^(NG|RG)") ~ "grass"
      )
  ) %>% 
  group_by(grass_crop) %>% 
  ggplot(aes(x = perc_silt, fill = grass_crop)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous() + 
  theme_bw()
```


```{r}
texture_stats %>% 
  group_by(grass_crop) %>% 
  ggplot(aes(x = perc_clay, fill = grass_crop)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous() + 
  theme_bw()
```

```{r}
texture_stats %>% 
  group_by(grass_crop) %>% 
  summarise(
    normality_p_val_sand = shapiro.test(perc_sand)$p.value,
    normality_p_val_silt = shapiro.test(perc_silt)$p.value,
    normality_p_val_clay = shapiro.test(perc_clay)$p.value
  )

```
```{r}
leveneTest(perc_sand ~ grass_crop, data = texture_stats) #unequal variances
leveneTest(perc_silt ~ grass_crop, data = texture_stats) #equal variances
leveneTest(perc_clay ~ grass_crop, data = texture_stats) #unequal variances
```

Wilcoxon Rank Sum Test
```{r}
wilcox.test(
  texture_stats$perc_sand ~ texture_stats$grass_crop, 
  paired = FALSE
)

wilcox.test(
  texture_stats$perc_silt ~ texture_stats$grass_crop, 
  paired = FALSE
)

wilcox.test(
  texture_stats$perc_clay ~ texture_stats$grass_crop, 
  paired = FALSE
)
```

### DOC means 
```{r}

doc %>% 
  group_by(disturbed) %>% 
  summarise(
    mean_doc = mean(blank_corrected_mg_l, na.rm = TRUE),
    se_doc = sd(blank_corrected_mg_l, na.rm = TRUE) / sqrt(n())
  ) %>% 
  knitr::kable()
  

```

### DOC Stats
```{r}
doc %>% 
  ggplot(aes(x = blank_corrected_mg_l, fill = disturbed)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous(limits = c(0, 250))

```


```{r}

doc %>% 
  group_by(disturbed) %>% 
  summarise(
    normality_p = shapiro.test(blank_corrected_mg_l)$p.value
  )

```
The data are approximately normally distributed (alpha = 0.05). 


```{r}
leveneTest(
  blank_corrected_mg_l ~ disturbed, 
  data = doc %>% 
    filter(!str_detect(sample_name, "Blank")) %>% 
    mutate(
      disturbed = if_else(str_detect(sample_name, "-D"), TRUE, FALSE)
    ) 
)

```
However, the variances are unequal. 


```{r}
t.test(
  x = doc %>% filter(disturbed == TRUE) %>% pull(blank_corrected_mg_l),
  y = doc %>% filter(disturbed == FALSE) %>% pull(blank_corrected_mg_l),
  var.equal = FALSE
  )
```


### Figure S5: DOC Boxplots
```{r}

doc %>% 
  filter(!str_detect(sample_name, "Blank")) %>% 
  mutate(
    disturbed = if_else(str_detect(sample_name, "-D"), TRUE, FALSE)
  ) %>% 
  ggplot(aes(x = disturbed, y = blank_corrected_mg_l, fill = disturbed)) + 
  geom_boxplot(color = "black") + 
  scale_fill_manual(values = c("gray60", "grey12")) + 
  scale_x_discrete(
    labels = c("Undisturbed\nGrassland", "Disturbed\nGrassland")
  ) + 
  labs(
    x = NULL,
    y = "DOC mg / L"
  ) + 
  theme_classic() + 
  theme(
    legend.position = "none"
  )

```





