---
title: "Ancillary Analyses"
author: "Emily Lacroix"
date: "4/16/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 300)
```


## Set-Up
### Load libraries
```{r,  results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(car)

conflicted::conflict_prefer("recode", "dplyr")
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
```

### Designate file paths
```{r}

data_file <- "~/Desktop/TempeCell_MasterSpreadsheet.xlsx"

```

### Figure labels

```{r}

field_labels <- 
  c(
    "RG-3" = "Riparian Grassland",
    "RG-2" = "Riparian Grassland",
    "RG-1" = "Riparian Grassland",
    "RG" = "Riparian Grassland",
    "3-8" = "Wheat/Cover",
    "1-6" = "Wheat/Cover",
    "7-3" = "Wheat/Cover",
    "4-9" = "Wheat/Fallow",
    "1-7" = "Wheat/Fallow",
    "8-3" = "Wheat/Fallow",
    "8-6" = "Native Grassland",
    "3-6" = "Native Grassland",
    "7-1" = "Native Grassland"
  )

```


## Import Data

### TC and TN data
```{r}
tc <- 
  data_file %>% 
  read_excel(sheet = "TotalCarbon", na = "NA") 

```

### pH data
```{r}
ph <- 
  data_file %>% 
  read_excel(sheet = "pH")

```

### XRF (total Fe & Mn) data
```{r}
xrf <-
  data_file %>% 
  read_excel(sheet = "XRF")

```

### Gravimetric data
```{r}
grav_data <- 
  read_xlsx(
    data_file, 
    sheet = "GravimetricData",
    na = "NA"
  ) 

```

### Texture data
```{r}
texture <-
  data_file %>% 
  read_excel(sheet = "Texture")

```

### DOC data
```{r}
doc <-
  data_file %>% 
  read_excel(sheet = "DissolvedOrganicCarbon") %>% 
  select(-comments) %>% 
  filter(!str_detect(sample_name, "Blank")) %>% 
  mutate(
    disturbed = if_else(str_detect(sample_name, "-D"), TRUE, FALSE)
  )

```

## Figures & Analysis

### Table 1: % OC and TN 
```{r}
tc %>% 
  mutate_at(
    vars(field),
    ~ recode(., !!!field_labels)
  ) %>% 
  mutate(
    grass_crop = 
      if_else(
        field %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  ) %>% 
  group_by(grass_crop) %>% 
  summarise_at(
    vars(perc_c, perc_n), 
    list(
      mean = ~mean(., na.rm = TRUE), 
      se = ~sd(., na.rm = TRUE) / sqrt(n())
    )
  ) %>% 
  select(grass_crop, perc_c_mean, perc_c_se, perc_n_mean, perc_n_se) %>% 
  knitr::kable()

```

### OC & TN Stats
```{r}
tc_stats <- 
  tc %>% 
  mutate_at(
    vars(field),
    ~ recode(., !!!field_labels)
  ) %>% 
  mutate(
    grass_crop = 
      if_else(
        field %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  )
  
```


```{r}

tc_stats %>% 
  group_by(grass_crop) %>% 
  summarise(
    normality_shapiro_p_oc = shapiro.test(perc_c)$p.value,
    normality_shapiro_p_tn = shapiro.test(perc_n)$p.value,
  )

```
The grassland data are not normally distributed. Checking for equal variances. 

```{r}
leveneTest(perc_c ~ grass_crop, data = tc_stats)
leveneTest(perc_n ~ grass_crop, data = tc_stats)
```

### Table 1: pH
```{r}

ph %>% 
  mutate(
    grass_crop = 
      if_else(
        crop %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  ) %>% 
  group_by(grass_crop) %>% 
  summarise_at(
    vars(pH), 
    list(
      mean = ~mean(., na.rm = TRUE), 
      se = ~sd(., na.rm = TRUE) / sqrt(n())
    )
  ) %>% 
  knitr::kable()
  
```

### pH Stats
```{r}
ph_stats <- 
  ph %>% 
  mutate(
    grass_crop = 
      if_else(
        crop %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  ) 
```


```{r}
ph_stats %>% 
  group_by(grass_crop) %>% 
  ggplot(aes(x = pH, fill = grass_crop)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous() + 
  theme_bw()

```

```{r}
ph_stats %>% 
  group_by(grass_crop) %>% 
  summarise(
    normality_p_val = shapiro.test(pH)$p.value
  )

```
The crop data are non-normal

```{r}
leveneTest(pH ~ grass_crop, data = ph_stats)
```
The variances are equal. Use Wilcoxon signed-rank test. 

```{r}
pairwise.wilcox.test(
  x = ph_stats$pH, 
  g = ph_stats$grass_crop, 
  p.adjust.method = "none", 
  paired = FALSE
)
```


### Table 1: Fe and Mn Content

For these data, we assume that soils across the grassland and cropland treatments have approximately equal elemental composition
```{r}
xrf %>% 
  summarise_at(
    vars(mn_ug_per_g, fe_ug_per_g), 
    list(
      mean = ~mean(., na.rm = TRUE),
      se =  ~sd(., na.rm = TRUE) / sqrt(n())
    )
  ) %>% 
  mutate(
    perc_fe = fe_ug_per_g_mean / 10^6 * 100, #unit converison
    se_perc_fe = fe_ug_per_g_se / 10^6 * 100, # unit conversion 
    mn_mg_kg = mn_ug_per_g_mean, # same values after unit conversion
    se_mn_mg_kg = mn_ug_per_g_se #same values after unit conversion
  ) %>% 
  select(perc_fe, se_perc_fe, mn_mg_kg, se_mn_mg_kg) %>% 
  knitr::kable()
```

### Table 1: Bulk Density
This focuses only on field cores, not disturbed cores as that bulk density is contrived. 
```{r}
grav_data %>% 
  mutate_at(
    vars(field),
    ~ recode(., !!!field_labels)
  ) %>% 
  filter(disturbed == FALSE) %>% 
  mutate(
    grass_crop = 
      if_else(
        field %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  ) %>% 
  group_by(grass_crop) %>% 
  summarise_at(
    vars(bulk_dens), 
    list(
      mean_bulk_dens = ~mean(., na.rm = TRUE), 
      se_bulk_dens = ~sd(., na.rm = TRUE) / sqrt(n())
    )
  ) %>% 
  knitr::kable()

```

### Table S2: Bulk Density by Field
This focuses only on field cores, not disturbed cores as that bulk density is contrived. 
```{r}
grav_data %>% 
  filter(disturbed == FALSE) %>% 
  group_by(field) %>% 
  summarise_at(
    vars(bulk_dens), 
    list(
      mean_bulk_dens = ~round(mean(., na.rm = TRUE), digits = 2), 
      se_bulk_dens = ~round(sd(., na.rm = TRUE) / sqrt(n()), digits = 2)
    )
  ) %>% 
  mutate(
    field_type = recode(field, !!!field_labels),
    grass_crop = 
      if_else(
        field_type %in% c("Riparian Grassland", "Native Grassland"), 
        "grass", 
        "crop"
      )
  ) %>% 
  select(grass_crop, field_type, field, mean_bulk_dens, se_bulk_dens) %>% 
  arrange(grass_crop, field_type, field) %>% 
  knitr::kable()

```


### Table 1: Texture
```{r}
texture %>% 
  mutate(
    grass_crop = 
      case_when(
        str_detect(sample, "^(WF|WC|CW)") ~ "crop",
        str_detect(sample, "^(NG|RG)") ~ "grass"
      )
  ) %>% 
  group_by(grass_crop) %>% 
  summarise_at(
    vars(perc_sand, perc_silt, perc_clay), 
    list(
      mean = ~mean(., na.rm = TRUE),
      se = ~sd(., na.rm = TRUE) / sqrt(n())
    )
  ) %>% 
  select(
    grass_crop,
    perc_sand_mean,
    perc_sand_se,
    perc_silt_mean,
    perc_silt_se,
    perc_clay_mean,
    perc_clay_se
  ) %>% 
  knitr::kable()

```


### DOC means 
```{r}

doc %>% 
  group_by(disturbed) %>% 
  summarise(mean_doc = mean(blank_corrected_mg_l, na.rm = TRUE)) %>% 
  knitr::kable()
  

```

### DOC Stats
```{r}
doc %>% 
  ggplot(aes(x = blank_corrected_mg_l, fill = disturbed)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous(limits = c(0, 250))

```


```{r}

doc %>% 
  group_by(disturbed) %>% 
  summarise(
    normality_p = shapiro.test(blank_corrected_mg_l)$p.value
  )

```
The data are approximately normally distributed (alpha = 0.05). 


```{r}
leveneTest(
  blank_corrected_mg_l ~ disturbed, 
  data = doc %>% 
    filter(!str_detect(sample_name, "Blank")) %>% 
    mutate(
      disturbed = if_else(str_detect(sample_name, "-D"), TRUE, FALSE)
    ) 
)

```
However, the variances are unequal. 


```{r}
t.test(
  x = doc %>% filter(disturbed == TRUE) %>% pull(blank_corrected_mg_l),
  y = doc %>% filter(disturbed == FALSE) %>% pull(blank_corrected_mg_l),
  var.equal = FALSE
  )
```


### Figure S5: DOC Boxplots
```{r}

doc %>% 
  filter(!str_detect(sample_name, "Blank")) %>% 
  mutate(
    disturbed = if_else(str_detect(sample_name, "-D"), TRUE, FALSE)
  ) %>% 
  ggplot(aes(x = disturbed, y = blank_corrected_mg_l, fill = disturbed)) + 
  geom_boxplot(color = "black") + 
  scale_fill_manual(values = c("gray60", "grey12")) + 
  scale_x_discrete(
    labels = c("Undisturbed\nGrassland", "Disturbed\nGrassland")
  ) + 
  labs(
    x = NULL,
    y = "DOC mg / L"
  ) + 
  theme_classic() + 
  theme(
    legend.position = "none"
  )

```





