---
title: "Pore Volume Analysis & Figures"
author: "Emily Lacroix"
date: "8/21/2020"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, dpi = 300)

```

## Set-Up

### Load libraries
```{r, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(car)

conflicted::conflict_prefer("recode", "dplyr")
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
```

### Designate filepaths
```{r Input files}

data_file <- "~/Desktop/TempeCell_MasterSpreadsheet.xlsx"

```

###  Figure labels
```{r Field Key, warning=FALSE}
#Field key
fields <- 
  tribble(
    ~crop_type, ~fields,
    "Native Grassland", c("7-1", "8-6", "3-6"),
    "Riparian Grassland", c("RG-1", "RG-2", "RG-3"),
    "Wheat/Cover", c("1-6", "3-8", "7-3"),
    "Wheat/Fallow", c("1-7", "4-9", "8-3"),
  ) %>% 
  unnest()
```


## Import & Clean Data

### Import
```{r Gravimetric Data}
grav_data <- 
  read_xlsx(
    data_file, 
    sheet = "GravimetricData",
    na = "NA"
  )
```

### Clean-up data

Read in all of the moisture data with the following filtering:

* remove Fe/Mn cores (we have no extraction data on these)
* Remove any cores that lost >= 5 g of moist soil under filter removal after extraction
* Retain all cores with > 95% WFPS (effectively saturated)
* Retain cores for which water recovered through pore water extraction and/or final oven drying (mw_resid) is 90-110% of the total soil pore volume as calculated by bulk density; assumed particle density of 2.65 g/cm3

```{r}
filtered_grav_data <- 
  grav_data %>% 
  filter(!str_detect(sample, "Fe")) %>%
  mutate(
    mw_resid =
      (extracted_soil + amount_removed - filter) -
      (dry_mass + ring_g + dry_boat_g),
    bulk_dens = dry_mass / ring_vol_cm3,
    porosity = 1 - (bulk_dens / 2.65),
    soil_pore_vol = porosity * ring_vol_cm3,
    alpha_100 = mw_100 / soil_pore_vol,
    alpha_300 = mw_300 / soil_pore_vol,
    alpha_500 = mw_500 / soil_pore_vol,
    alpha_less_than_500 = mw_resid / soil_pore_vol
  ) %>%
  drop_na(mw_100, mw_300, mw_500) %>%
  filter(wfps > 0.95) %>%
  mutate(
    error =
      (mw_100 + mw_300 + mw_500 + mw_resid - soil_pore_vol) / soil_pore_vol
  )  %>%
  filter(abs(error) <= 0.10) %>%
  left_join(fields, by = c("field" = "fields")) %>%
  mutate(
    tillage =
      case_when(
        crop_type %in% c("Wheat/Fallow", "Wheat/Cover") ~ "Tilled Cropland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") &
          disturbed == FALSE ~ "Undisturbed Grassland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") &
          disturbed == TRUE ~ "Disturbed Grassland"
      )
  ) %>% 
  filter(amount_removed < 5)


```


## Statistics 

### Density Plots
Density plots of each of the proportion pore volumes for the cores (after quality filtering)
```{r}

filtered_grav_data %>% 
  ggplot(aes(x = alpha_100, fill = tillage)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous(limits = c(0,0.8)) + 
  theme_bw()


filtered_grav_data %>% 
  ggplot(aes(x = alpha_300, fill = tillage)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous(limits = c(0,0.2)) + 
  theme_bw()

filtered_grav_data %>% 
  ggplot(aes(x = alpha_500, fill = tillage)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous(limits = c(0,0.1)) + 
  theme_bw()

filtered_grav_data %>% 
  ggplot(aes(x = alpha_less_than_500, fill = tillage)) + 
  geom_density(alpha = 0.5) + 
  scale_x_continuous(limits = c(0,1)) + 
  theme_bw()

```


### Test for normality with the Shapiro test 
```{r}

filtered_grav_data %>% 
  group_by(tillage) %>%
  summarise_at(vars(contains("alpha")), ~shapiro.test(.)$p.value) %>% 
  rename_at(vars(contains("alpha")), ~ str_c(., "_shapiro_p")) %>% 
  knitr::kable()

```

Disturbed grassland pore volumes are NOT normal (for >3um, 1-3 um kPa, and < 0.6 um)

### Check for equal variances with Levene's test

```{r, warning=FALSE}

leveneTest(alpha_100 ~ tillage, data = filtered_grav_data)
leveneTest(alpha_300 ~ tillage, data = filtered_grav_data)
leveneTest(alpha_500 ~ tillage, data = filtered_grav_data)
leveneTest(alpha_less_than_500 ~ tillage, data = filtered_grav_data)


```
All of the population variances are equal. Now we can use a pairwise Wilcox test to look for differences.

### Table S5: Pairwise Wilcoxon rank sum test for percent pore volumes
```{r}

pairwise.wilcox.test(
  x = filtered_grav_data$alpha_100, 
  g = filtered_grav_data$tillage, 
  p.adjust.method = "BH", 
  paired = FALSE
)

pairwise.wilcox.test(
  x = filtered_grav_data$alpha_300, 
  g = filtered_grav_data$tillage, 
  p.adjust.method = "BH", 
  paired = FALSE
)

pairwise.wilcox.test(
  x = filtered_grav_data$alpha_500, 
  g = filtered_grav_data$tillage, 
  p.adjust.method = "BH", 
  paired = FALSE
)

pairwise.wilcox.test(
  x = filtered_grav_data$alpha_less_than_500, 
  g = filtered_grav_data$tillage, 
  p.adjust.method = "BH", 
  paired = FALSE
)

```



## Figures & Tables

### Section 3.2: Average pores < 0.6 micron
```{r}
filtered_grav_data %>% 
  group_by(tillage) %>%
  summarise_at(
    vars(contains("alpha")),
    list(
      mean = ~ 100 * mean(., na.rm = TRUE),
      se = ~ 100 * sd(., na.rm = TRUE)/sqrt(n())
    )
  ) %>%
  select(
    tillage,
    alpha_100_mean,
    alpha_100_se,
    alpha_300_mean,
    alpha_300_se,
    alpha_500_mean,
    alpha_500_se,
    alpha_less_than_500_mean,
    alpha_less_than_500_se
  ) %>% 
  summarise(mean(alpha_less_than_500_mean))
```


### Figure 1. Percent Pore Volume 
```{r}

filtered_grav_data %>% 
  group_by(tillage) %>%
  summarise_at(
    vars(contains("alpha")),
    list(
      mean = ~mean(., na.rm = TRUE),
      se = ~sd(., na.rm = TRUE)/sqrt(n())
    )
  ) %>%
  select(
    tillage,
    alpha_100_mean,
    alpha_100_se,
    alpha_300_mean,
    alpha_300_se,
    alpha_500_mean,
    alpha_500_se,
    alpha_less_than_500_mean,
    alpha_less_than_500_se
  ) %>% 
  pivot_longer(
    cols = 
      c(
        alpha_100_mean, 
        alpha_300_mean, 
        alpha_500_mean, 
        alpha_less_than_500_mean,
        alpha_100_se, 
        alpha_300_se, 
        alpha_500_se, 
        alpha_less_than_500_se
      ),
    names_to = "pore_size_value",
    values_to = "proportion"
  ) %>% 
  mutate(
    value = str_extract(pore_size_value, "mean|se"),
  ) %>% 
  mutate_at(
    vars(pore_size_value), 
    ~ str_replace(., "_mean|_se", "")
  ) %>% 
  pivot_wider(
    id_cols = c(tillage, pore_size_value),
    names_from = value,
    values_from = proportion
  ) %>% 
  mutate_at(
    vars(tillage),
    ~ factor(
      .,
      levels =
        c(
          "Undisturbed Grassland",
          "Tilled Cropland",
          "Disturbed Grassland"
        )
    )
  ) %>% 
  mutate_at(
    vars(pore_size_value),
    ~ factor(
      .,
      levels = c("alpha_100", "alpha_300", "alpha_500", "alpha_less_than_500"),
      ordered = TRUE,
      labels =
        c(
          expression(paste("> 3.0 ", mu, "m")),
          expression(paste("3.0 - 1.0 ", mu, "m")),
          expression(paste("1.0 - 0.6 ", mu, "m")),
          expression(paste("< 0.6 ", mu, "m"))
        )
    )
  ) %>% 
  ggplot(aes(y = mean, x = tillage, fill = tillage)) + 
  geom_col(color = "black") + 
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.4) + 
  scale_fill_manual(values = c("gray60", "grey39", "grey12")) + 
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = c(0.01,0.005, 0.01, 0.1)
  ) + 
  facet_grid(cols = vars(pore_size_value), labeller = label_parsed) + 
  labs(
    fill = NULL,
    x = NULL,
    y = "Percent of Pore Volume"
  ) + 
  theme_bw() + 
  theme(
    rect = element_rect(fill = "transparent"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    axis.text.y = element_text(size = 14, color = "black"),
    axis.title.y = element_text(size = 16), 
    legend.text = element_text(size = 12),
    strip.text.x = element_text(size = 13, face = "bold"),
    panel.background = element_rect(fill = "transparent"), 
    plot.background = element_rect(fill = "transparent", color = NA), 
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
  ) 

```

### Table S7: Total Water Extracted by Moisture and Disturbance
```{r}
  grav_data %>% 
  filter(!str_detect(sample, "Fe")) %>%
  drop_na(mw_100, mw_300, mw_500) %>%
  mutate(
    water_extracted = mw_100 + mw_300 + mw_500,
    moisture = if_else(wfps < 0.8, "low", "high")
  )  %>%
  left_join(fields, by = c("field" = "fields")) %>%
  mutate(
    tillage =
      case_when(
        crop_type %in% c("Wheat/Fallow", "Wheat/Cover") ~ "Tilled Cropland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") &
          disturbed == FALSE ~ "Undisturbed Grassland",
        crop_type %in% c("Native Grassland", "Riparian Grassland") &
          disturbed == TRUE ~ "Disturbed Grassland"
      )
  ) %>% 
  group_by(moisture, tillage) %>% 
  summarise(
    mean_total = mean(water_extracted),
    se_total = sd(water_extracted) / sqrt(n()),
    mean_100 = mean(mw_100),
    se_100 = sd(mw_100) / sqrt(n()),
    mean_300 = mean(mw_300),
    se_300 = sd(mw_300) / sqrt(n()),
    mean_500 = mean(mw_500),
    se_500 = sd(mw_500) / sqrt(n()),
  ) %>% 
  arrange(desc(tillage), desc(moisture)) 
  

```


